---
title: "AI Deserts - Overlap with Infertility-Associated Loci"
author: "Maegwin Bonar"
date: "2025-07-11"
output:
  html_document:
    highlight: breezedark
    css:
      - "styles.css"
    toc: yes
    toc_float: no
    toc_depth: 5
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval                        = FALSE,
  echo                        = TRUE,
  cache                       = TRUE, # tidy = TRUE,
  class.output                = "bg-success"
)
```

### Introduction

We obtained Neanderthal introgression deserts from Sankararaman et al. (2016), and used a hypergeometric test in R to test over-representation of genes containing infertility-associated loci within introgression deserts.


### AI deserts

Use deserts from Sankararaman et al. 2016
Input: Sankararaman_2016_NEA_deserts.bed

Intersect Sankararaman et al. 2016 files with genecode bed files
```{bash}

bedtools intersect -wa -wb -a Sankararaman_2016_NEA_deserts.bed -b gencode.v38lift37.annotation.codingGENES_sorted.bed > Sankararaman_2016_NEA_deserts_intersect_genes.bed

```

Intersect with infertility genes. Only intersected with regions from Deserts from Sankararaman et al. 2016 
```{bash}

bedtools intersect -wa -wb -a Sankararaman_2016_NEA_deserts.bed -b infertility_genes.bed > Sankararaman_2016_NEA_intersect_infertilitygenes_2025.bed

```

In this case we will check whether genes associated with infertility are over-represented in genes overlapping with AI deserts
```{r}

## get the values for each parameter
# we need `genes`, `gene_set`, and `universe`

### genes - infertility genes
genes <- infert_genes_all[1]

### gene set - all genes in AI deserts NEA
gene_set_nea<- read.table("Sankararaman_2016_NEA_deserts_intersect_genes.bed", header = FALSE)
colnames(gene_set_nea) <- c("chr", "start","end","chr.y", "start.y","end.y","ensembl", "gene")

## remove duplicates
gene_set_nea <- gene_set_nea %>%
  distinct(gene)

### universe - all genes
universe<- read.table("gencode.v38lift37.annotation.codingGENES_sorted.bed", header = FALSE)
colnames(universe) <- c("chr", "start","end","ensembl", "gene")

universe<-universe %>%
  select(gene) %>%
  distinct(gene)

#### Test
ora_ai_nea <- ora_single(genes = genes$gene, gene_set = gene_set_nea$gene, universe = universe$gene)
ura_ai_nea <- ura_single(genes = genes$gene, gene_set = gene_set_nea$gene, universe = universe$gene)

### as a check, the intersection of genes and gene_set should be x
# remove duplicates and take only gene column
ai_infert_genes <- read.table("Sankararaman_2016_NEA_deserts_intersect_infertilitygenes.bed", header = FALSE)
colnames(ai_infert_genes) <- c("chr", "start","end","chr.y", "start.y","end.y","ensembl", "gene")

x <- ai_infert_genes %>%
  select(gene) %>%
  distinct(gene)

x_check = intersect(genes$gene, gene_set_nea$gene)

```
