---
title: "Genetic Contraint Analysis"
author: "Maegwin Bonar"
date: "2025-02-12"
output:
  html_document:
    highlight: breezedark
    css:
      - "styles.css"
    toc: yes
    toc_float: no
    toc_depth: 5
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval                        = FALSE,
  echo                        = TRUE,
  cache                       = TRUE, # tidy = TRUE,
  class.output                = "bg-success"
)
knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)
```

### Introduction

We expect that genes associated with infertility will be highly constrained

Assess constraint using three metrics: pLI, LOEUF, and z-score using table to constraints from gnomeAD


### Libraries
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggpubr)
```


### 1. Read in the data
```{r}
## read in the table of constraints from gnomeAD
gnomeAD_const <- read.table("/gnomAD/v4.0.0/gnomad.v4.1.constraint_metrics.tsv", header = TRUE)
head(gnomeAD_const)

## read in table of genes of interest (infertility)
genes_infert <- read.table("infertilityGeneList_unique.txt", header = FALSE)
colnames(genes_infert) <- c("gene", "type")
```

### 2. Genetic constraints on genes assoicated with infertility

Get the constraint metrics for each gene of interest
```{r}
# left join with genes of interest
genes_const <- na.omit(left_join(genes_infert, gnomeAD_const, by = "gene"))

mn_pLI <- mean(genes_const$lof.pLI)
mn_LOEUF <- mean(genes_const$lof.oe_ci.upper)

med_pLI <- median(genes_const$lof.pLI)
min(genes_const$lof.pLI)
max(genes_const$lof.pLI)
med_LOEUF <- median(genes_const$lof.oe_ci.upper)
min(genes_const$lof.oe_ci.upper)
max(genes_const$lof.oe_ci.upper)

mn_zscore <- mean(genes_const$lof.z_score)
med_zscore <- median(genes_const$lof.z_score)
min(genes_const$lof.z_score)
max(genes_const$lof.z_score)

genes_results <- data.frame(mn_pLI, med_pLI, mn_LOEUF, med_LOEUF, mn_zscore, med_zscore)

### Plot the distribution of metrics
# lofpLI
a<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.pLI), binwidth = 0.1)+
  #ylim(0,1)+
  #geom_vline(xintercept = mn_pLI, colour = "blue")+
  geom_vline(xintercept = med_pLI, colour = "#C05227", linewidth = 1.2)+
  #facet_wrap(~type, nrow = 2)+
  theme_pubr()
  
# LOEUF
b<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.oe_ci.upper), binwidth = 0.1)+
  #geom_vline(xintercept = mn_LOEUF, colour = "blue")+
  geom_vline(xintercept = med_LOEUF, colour = "#C05227", linewidth = 1)+
  #facet_wrap(~type, nrow = 2)+
  theme_pubr()

# zscore
c<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.z_score), binwidth = 1)+
  #geom_vline(aes(xintercept = mn_zscore, colour = "mean"))+
  geom_vline(aes(xintercept = med_zscore), colour = "#C05227", linewidth = 1)+
  #scale_colour_manual(name = "statistics", values = c(median = "#1B3F74", mean = "blue"))+
  #facet_wrap(~type, nrow = 2)+
  theme_pubr()


d<-ggarrange(a, b, c, nrow = 1)

ggsave("plots/genes_constraint.pdf", d, width = 10, height = 4)
#ggsave("plots/genes_constraint2.pdf", d, width = 10, height = 8)

# Compute median pLI per group
medians_pLI <- genes_const %>%
  group_by(type) %>%
  summarize(med_pLI = median(lof.pLI, na.rm = TRUE))

medians_LOEUF <- genes_const %>%
  group_by(type) %>%
  summarize(med_LOEUF = median(lof.oe_ci.upper, na.rm = TRUE))

medians_zscore <- genes_const %>%
  group_by(type) %>%
  summarize(med_zscore = median(lof.z_score, na.rm = TRUE))

# Plot with facet-specific vlines
ggplot(data = genes_const) +
  geom_histogram(aes(x = lof.pLI), binwidth = 0.1) +
  geom_vline(data = medians, aes(xintercept = med_pLI), 
             colour = "#C05227", linewidth = 1.2) +
  facet_wrap(~type, nrow = 2) +
  theme_pubr() +
  theme(axis.title.y = element_blank())

# lofpLI
a<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.pLI), binwidth = 0.1)+
  #ylim(0,1)+
  #geom_vline(xintercept = mn_pLI, colour = "blue")+
  geom_vline(data = medians_pLI, 
             aes(xintercept = med_pLI), colour = "#C05227", linewidth = 1.2)+
  facet_wrap(~type, nrow = 2)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )
  
# LOEUF
b<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.oe_ci.upper), binwidth = 0.1)+
  #geom_vline(xintercept = mn_LOEUF, colour = "blue")+
  geom_vline(data=medians_LOEUF, aes(xintercept = med_LOEUF), colour = "#C05227", linewidth = 1)+
  facet_wrap(~type, nrow = 2)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )

# zscore
c<-ggplot(data = genes_const)+
  geom_histogram(aes(x=lof.z_score), binwidth = 1)+
  #geom_vline(aes(xintercept = mn_zscore, colour = "mean"))+
  geom_vline(data=medians_zscore, aes(xintercept = med_zscore), colour = "#C05227", linewidth = 1)+
  #scale_colour_manual(name = "statistics", values = c(median = "#1B3F74", mean = "blue"))+
  facet_wrap(~type, nrow = 2)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )


d<-ggarrange(a, b, c, nrow = 1)
plot_genes <- annotate_figure(d, top = text_grob("Constraint metrics for 24 infertility genes", 
               color = "black", face = "bold", size = 14))

ggsave("plots/genes_constraint_m_f.pdf", plot_genes, width = 10, height = 6)
ggsave("plots/genes_constraint_m_f.png", plot_genes, width = 10, height = 6)

```


### 3. Compare to background distribution

Sample from whole gene set (including infertility genes) the same number of genes (24)
Generate distribution of constraint metrics for sampled set
Repeat 10,000 times to generate confidence intervals around the distribution

```{r}

#get the list of genes from the gnomeAD data
gnomeAD_clean <- na.omit(gnomeAD_const)
gnomeAD_clean_genes <- unique(gnomeAD_const$gene)

ggplot(data = gnomeAD_clean)+
  geom_histogram(aes(x=lof.pLI))

ggplot(data = gnomeAD_clean)+
  geom_histogram(aes(x=lof.z_score))

# number of iterations
num_iterations <- 10000

# results list
results <- data.frame(mn_pLI = integer(num_iterations), 
                      med_pLI = integer(num_iterations),
                      mn_LOEUF = integer(num_iterations),
                      med_LOEUF = integer(num_iterations),
                      mn_zscore = integer(num_iterations),
                      med_zscore = integer(num_iterations))

# Initialize the progress bar
pb <- txtProgressBar(min = 0, max = num_iterations, style = 3)

# Loop through each iteration
for (i in 1:num_iterations) {
  # Randomly sample individuals from the dataframe
  sampled_genes <- sample(gnomeAD_clean_genes, 25, replace = FALSE)

  # get genetic constraints for this set of genes
  sample_genes_const <- na.omit(gnomeAD_const[ gnomeAD_const$gene %in% sampled_genes,])
  
  # calculate summary metrics
  mn_pLI <- mean(sample_genes_const$lof.pLI)
  mn_LOEUF <- mean(sample_genes_const$lof.oe_ci.upper)

  med_pLI <- median(sample_genes_const$lof.pLI)
  med_LOEUF <- median(sample_genes_const$lof.oe_ci.upper)
  
  mn_zscore <- mean(sample_genes_const$lof.z_score)
  med_zscore <- median(sample_genes_const$lof.z_score)

  # assign genetic constraint metrics to results list
  results[i,] <- c(mn_pLI, med_pLI, mn_LOEUF, med_LOEUF, mn_zscore, med_zscore)
  # Update the progress bar
  setTxtProgressBar(pb, i)
}

close(pb)  # Close the progress bar after the loop

```

Plot the background set compared to infertility genes

```{r}
### Plot the distribution of metrics

### Mean
# lofpLI
p1<-ggplot(data = results)+
  geom_histogram(aes(x=mn_pLI), binwidth = 0.02)+
  geom_vline(xintercept = genes_results$mn_pLI, colour = "#C05227")+
  theme_bw()
  
# LOEUF
p2<-ggplot(data = results)+
  geom_histogram(aes(x=mn_LOEUF), binwidth = 0.02)+
  geom_vline(xintercept = genes_results$mn_LOEUF, colour = "#C05227")+
  theme_bw()

# zscore
p3<-ggplot(data = results)+
  geom_histogram(aes(x=mn_zscore), binwidth = 0.1)+
  geom_vline(xintercept = genes_results$mn_zscore, colour = "#C05227")+
  theme_bw()

plot_mean<-ggarrange(p1, p2, p3, nrow = 1)
plot_mean2 <- annotate_figure(plot_mean, top = text_grob("Mean genetic constraints sampled from gnomeAD genes", 
               color = "black", face = "bold", size = 14))

ggsave("plots/sample_genes_constraint_mean.pdf", plot_mean2, width = 10, height = 4)


### Median
# lofpLI
p4<-ggplot(data = results)+
  geom_histogram(aes(x=med_pLI), binwidth = 0.05)+
  geom_vline(xintercept = genes_results$med_pLI, colour = "#C05227", linewidth = 1)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )
  
# LOEUF
p5<-ggplot(data = results)+
  geom_histogram(aes(x=med_LOEUF), binwidth = 0.05)+
  geom_vline(xintercept = genes_results$med_LOEUF, colour = "#C05227", linewidth = 1)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )

# zscore
p6<-ggplot(data = results)+
  geom_histogram(aes(x=med_zscore), binwidth = 0.1)+
  geom_vline(xintercept = genes_results$med_zscore, colour = "#C05227", linewidth = 1)+
  theme_pubr()+
  theme(
    axis.title.y = element_blank()
  )

plot_med<-ggarrange(p4, p5, p6, nrow = 1)
plot_med2 <- annotate_figure(plot_med, top = text_grob("Median genetic constraints sampled from gnomeAD genes", 
               color = "black", face = "bold", size = 14))

plot_fig1b <- ggarrange(a, b, c, p4, p5, p6, ncol = 3, nrow = 2)

ggsave("plots/sample_genes_constraint_med.png", plot_med2, width = 10, height = 4)
ggsave("plots/sample_genes_constraint_med.pdf", plot_med, width = 10.5, height = 4)

ggsave("plots/constraint_fig1b.pdf", plot_fig1b, width = 11, height = 5)

```


MonteCarlo statistical test
Check effect size

```{r}
## for pLI
## what proportion of the distribution =< my value
p_pLI <- mean(results$med_pLI >= genes_results$med_pLI)
p_pLI
## for LOEUF
## what proportion of the distribution >= my value
p_LOEUF <- mean(results$med_LOEUF <= genes_results$med_LOEUF)
p_LOEUF
## for zscore
## what proportion of the distribution >= my value
p_zscore <- mean(results$med_zscore >= genes_results$med_zscore)
p_zscore
#save.image("constraints.RData")

#load("constraints.RData")
```


